Zambia Potential Yield Grids

### 2016-01-28 Rmd/all-inputs.Rmd
Bunch of stuff, up through creation of figures for presentation, in demo-figures.Rmd

2. For running model:
    + Prioritization (P) and impact (I)
        + `ZA-carbon.tif` - Carbon impacts, to data.table (P|I)
        + `bdprior.tif` - biodiversity prioritization grid (P)
        + `ZA-cost.tif` - cost distance
        + `maize|soybean_S|R|I_10km,tif` - variants thereof (P|I). Separate variants of these must be created prior to model runtime. A function will be creatd here to facilitate that.  
    + Impact only:
        + `ZA-vegremain.tif` - vegetation remaining grid, for assessing conversion impacts. 
    + Masking/allocation
        + `farmable-1k.tif` - slope-based farmable area mask
        + `water-1k.tif` - water-based farmable area mask
        + `ZA-remfrac.tif` - cropland and urban-based farmable area mask
3. For displaying model outputs: 
    + `cropland-1k.tif` - cropland mask

@@

p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_devel <- full_path(p_proj, "external/input_devel/cropland")
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_crop <- full_path(p_proj, "external/input_devel/cropland")
p_bd <- full_path(p_proj, "external/input_devel/biodiversity")
p_cost <- full_path(p_proj, "external/input_devel/roads/cost")

# Load in data
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))
cropland <- raster(full_path(p_crop, fname = "cropland-1k.tif"))
bd <- raster(full_path(p_bd, fname = "bdprior.tif"))
vegr <- raster(full_path(p_bd, fname = "ZA-vegremain.tif"))
cost <- raster(full_path(p_cost, fname = "zamcost.tif"))

@@

>>>>

2. For running model:
    + Prioritization (P) and impact (I)
        + `ZA-carbon.tif` - Carbon impacts, to data.table (P|I)
        + `bdprior.tif` - biodiversity prioritization grid (P)
        + `ZA-cost.tif` - cost distance
        + `maize|soybean_S|R|I_10km,tif` - variants thereof (P|I). Separate variants of these must be created prior to model runtime. A function will be creatd here to facilitate that.  
    + Impact only:
        + `ZA-bd.tif` - vegetation remaining grid, for assessing conversion impacts. 
        + `intactness.tif` - measure of intactness
        + `ZA-pas.tif` - protected areas, to see how many forest reserves get hit
    + Masking/allocation
        + `farmable-1k.tif` - slope-based farmable area mask
        + `water-1k.tif` - water-based farmable area mask
        + `ZA-remfrac.tif` - cropland and urban-based farmable area mask
3. For displaying model outputs: 
    + `cropland-1k.tif` - cropland mask

@@

p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_devel <- full_path(p_proj, "external/input_devel/cropland")
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_crop <- full_path(p_proj, "external/input_devel/cropland")
p_bd <- full_path(p_proj, "external/input_devel/biodiversity")
p_cost <- full_path(p_proj, "external/input_devel/roads/cost")
p_yield <- full_path(p_proj, "external/input_devel/yield")
p_h20 <- full_path(p_proj, "external/input_devel/water")
p_terr <- full_path(p_proj, "external/input_devel/terrain")
p_carb <- full_path(p_proj, "external/input_devel/carbon")
p_production <- full_path(p_proj, "external/data/production")

# Load in data
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))
cropland <- raster(full_path(p_crop, fname = "cropland-1k.tif"))
cons <- raster(full_path(p_bd, fname = "cons-priorities.tif"))
bd <- brick(full_path(p_bd, fname = "ZA-vegremain.tif"))
load(full_path(p_bd, fname = "ZA-vegremain.rda"))
names(bd) <- vegnames
carbon <- brick(full_path(p_carb, fname = "ZA-carbon.tif"))
names(carbon) <- c("veg", "soil")
cost <- raster(full_path(p_cost, fname = "ZA-cost.tif"))
pas <- raster(full_path(p_bd, "ZA-pas.tif"))
intact <- raster(full_path(p_bd, "intactness.tif"))
ynms <- dir(p_yield, pattern = "FAO.tif")  # FAO adjusted now
ynms <- sapply(c("S", "R", "I"), function(x) ynms[grep(x, ynms)])
ynms <- ynms[c(1, 3, 5, 2, 4, 6)]
ylds <- lapply(ynms, function(x) {
  b <- brick(full_path(p_yield, x))
  ylnms <- gsub("\\.tif", "\\.rda", x)
  load(full_path(p_yield, ylnms))
  names(b) <- layernms
  b
})
names(ylds) <- gsub("_10K|\\.tif|_FAO", "", ynms)

# mask materials
remfrac <- raster(full_path(p_bd, fname = "ZA-remfrac.tif"))
water <- raster(full_path(p_h20, fname = "water-1k.tif"))
farmable <- raster(full_path(p_terr, fname = "farmable-1k.tif"))

@@

## Create farmable area

```{r, eval=FALSE}
# merge together the three farmland available masks
avail <- remfrac - water
avail[avail < 0] <- 0
avail <- avail - (1 - farmable)
avail[avail < 0] <- 0
avail <- avail * (cons < 0.962)  # remove PAs entirely (1 if nps, 0.962 if GMAs)
names(avail) <- "convertible"
# plot(avail)
```

That creates the farmland availability mask that is needed.

## Stack and mask all grids
```{r, eval = FALSE}
# deal with yields
potmaize <- ylds$maize_R  # just taking rainfed commercial potential to start
w <- matrix(nrow = 5, ncol = 5, rep(1, 5 * 5))  # have to filter b/c artifacts
potmaizef <- focal(potmaize[[1]], w = w, fun = mean, na.rm = TRUE)
potsoy <- ylds$soybean_R  # just taking rainfed commercial potential to start
potsoy[potsoy > 5] <- 5 #round(cellStats(potsoy, mean), 1)  # really high fix
potsoyf <- focal(potsoy[[1]], w = w, fun = mean, na.rm = TRUE)
# potsoy <- mask(potsoyf, potsoy)  # crop yield

yield_fix <- function(x = r[[1]], w, na.rm = TRUE) {
  r <- focal(x, w = w, fun = mean, na.rm = na.rm)
  rmsk <- (!is.na(r) & is.na(x)) * r
  x[is.na(x)] <- 0
  rfix <- x + rmsk  # add back in edge bits
}

w <- matrix(nrow = 3, ncol = 3, rep(1, 9))
potsoyf2 <- yield_fix(potsoyf[[1]], w)
potmaizef2 <- yield_fix(potmaizef[[1]], w)
plot(potmaizef2)
# first get rid of all NA area in all grids
# start with yield grids
# yldmask <- calc(stack(ylds), sum)

# Function to disaggregate and reproject yield rasters
yld_disagg <- function(x, msk, fact = 10) {
  r <- disaggregate(x, fact = 10)
  r2 <- resample(x, r)
  r3 <- projectRaster(r2, msk)
  r4 <- mask(r3, msk)
  r4
}

ylds_1k <- lapply(list(potmaizef2, potsoyf2), function(x) {
  yld_disagg(x, msk = mgrid)
})
ylds_s <- stack(ylds_1k) 
names(ylds_s) <- c("maize", "soy")
ylds_s <- round(ylds_s, 1)  
# plot(ylds_s)

# yldmask <- calc(stack(potmaizef2, potsoyf2), sum)
# yldmask2 <- disaggregate(yldmask, fact = 10)
# yldmask3 <- resample(yldmask, yldmask2)
# yldmask4 <- projectRaster(yldmask3, mgrid)
# yldmask5 <- mask(yldmask4, mgrid)
# plot(yldmask5)

# now with other
carbonf <- stack(lapply(1:nlayers(carbon), function(x) {
  r <- carbon[[x]]
  r[is.na(r)] <- 0
  r
}))

# pas and intactness
intpa <- stack(intact, pas)

mgridc <- crop(mgrid, avail)
inlist <- list("convertible" = avail, "bd" = bd, "carbon" = carbonf, 
               "cost" = cost, "cons" = cons, "cropland" = cropland, 
               "intpa" = intpa, "p_yield" = ylds_s)
inlist <- lapply(inlist, function(x) crop(x, mgridc))
inlist2 <- lapply(inlist, function(x) calc(x, mean))
inlist3 <- calc(stack(inlist2), mean)
# inlist4 <- calc(stack(inlist3, crop(ylds_s, mgridc)), mean)

# filter NA values
# nafilt <- calc(stack(!is.na(inlist3), !is.na(crop(yldmask5, mgridc))), sum)
nafilt <- !is.na(inlist3)
# plot(nafilt)
# plot(!is.na(calc(vegr, mean)))
# plot(!is.na(crop(yldmask5, mgridc)), col = c("transparent", "red"), add = TRUE)
msk <- nafilt == maxValue(nafilt)  # freq(nafilt)
msk[msk == 0] <- NA
# plot(msk)

# updated mask
mgridf <- mask(mgridc, msk)
valinds <- which(!is.na(values(mgridf)))
outlist <- inlist[names(inlist) != "cropland"]

# write out new data.tables
### make sure convertible writes out column header name properly
cnames <- "ZA"  
onms <- names(outlist)
onms[onms == "cons"] <- "cons-priorities"
onms[onms == "p_yield"] <- "potential-yields"

for(i in 1:length(outlist)) { # i <- 1
  fnm <- full_path(p_production, paste0(cnames, "-", onms[i], ".csv"))
  DT <- as.data.table(outlist[[i]], xy = FALSE)[valinds, ]
  write.table(DT, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
}

# write out mask raster
base <- as.data.table(msk, xy = TRUE)
setnames(base, old = names(msk), new = "val")
base[, ind := 1:nrow(base)]
base <- base[!is.na(val)][, val := NULL]
fnm <- full_path(p_production, paste0(cnames, "-mask.csv"))
write.table(base, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
# plot(dt_to_raster(base, mgridf@crs))

# # data.table them
# cnames <- "ZA"  
# fnm <- full_path(p_production, paste0(cnames, "-potential-yields.csv"))
# DT <- as.data.table(ylds_s, xy = FALSE)[valinds, ]
# # DT[which(is.na(DT))]
# write.table(DT, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
```

# Copy over files to the correct directory
```{r, eval}
p_mod <- full_path(p_proj, "agroEcoTradeoff/external/data/dt/latest/")
p_master <- "~/Dropbox/projects/agroEcoTradeoff/external/data/dt/latest/"
fnms <- dir(p_production)
file.copy(dir(p_production, full.names = TRUE), full_path(p_mod, fnms),
          overwrite = TRUE)
file.copy(dir(p_production, full.names = TRUE), full_path(p_master, fnms),
          overwrite = TRUE)

## Yield grids
```{r, eval = FALSE}
# # Function to disaggregate and reproject yield rasters
# yld_disagg <- function(x, msk, fact = 10) {
#   r <- disaggregate(x, fact = 10)
#   r2 <- resample(x, r)
#   r3 <- projectRaster(r2, msk)
#   r4 <- mask(r3, msk)
#   r4
# }

# # Function to aggregate yields within a particular yield brick
# # stat <- "50"; cults <- paste0("9900", c("08", "09", "10"))
# yield_stat <- function(x, stat = "mu", mgmnt = "S", cults, fun = mean) {
#   repstr <- paste0(paste0(stat, "_", mgmnt), cults, collapse = "|")
#   cnms <- names(x)[grep(repstr, names(x))]
#   print(cnms)
#   calc(x[[cnms]], fun = fun)
# }
# rm(stat, mgmnt, cults, fun, x, cnms, repstr)

# Let's make an initial yield combination for model development that is
# is composed of: 
# maize
# 1. Subsistence management, all four cultivars, median yields, 50%
# 2. Rainfed commerical, hybrid cultivars, median yields, 40%
# 3. Irrigated maize, hybrid cultivars, median yields, 10%

# maize <- list("S" = yield_stat(x = ylds$maize_S.tif, "50", "S", 
#                                paste0("000", 1:4)), 
#               "R" = yield_stat(x = ylds$maize_R.tif, "50", "R", 
#                                paste0("000", 3:4)), 
#               "I" = yield_stat(x = ylds$maize_I.tif, "50", "I", 
#                                paste0("000", 3:4)))
# wgts <- c(0.5, 0.4, 0.1)
# maizewgt <- lapply(1:length(wgts), function(x) maize[[x]] * wgts[x])
# potmaize <- calc(stack(maizewgt), sum)  # potential maize yield

# wgts <- c(0.2, 0.7, 0.1)
# mznms <- names(ylds)[grep("maize", names(ylds))]
# # maizewgt <- lapply(1:length(wgts), function(x) ylds[[mznms[x]]] * wgts[x])
# # potmaize <- calc(stack(maizewgt), sum)  # potential maize yield
# potmaize <- ylds$maize_R  # just taking rainfed commercial potential to start
# plot(ylds$maize_I)

# Soybean
# 1. Subsistence management, all 3 cultivars, median yields, 20%
# 2. Rainfed commerical, all 3cultivars, median yields, 70%
# 3. Irrigated maize, all 3 cultivars, median yields, 10%
# soy <- list("S" = yield_stat(x = ylds$soybean_S.tif, "50", "S", 
#                              paste0("9900", c("08", "09", "10"))), 
#             "R" = yield_stat(x = ylds$soybean_R.tif, "50", "R", 
#                              paste0("9900", c("08", "09", "10"))), 
#             "I" = yield_stat(x = ylds$soybean_I.tif, "50", "I", 
#                              paste0("9900", c("08", "09", "10"))))
# wgts <- c(0.2, 0.7, 0.1)
# soywgt <- lapply(1:length(wgts), function(x) soy[[x]] * wgts[x])
# potsoy <- calc(stack(soywgt), sum)  # potential maize yield
# plot(potmaize); cellStats(potmaize, mean)
# plot(potsoy); cellStats(potsoy, mean)
# plot(potmaize)
# plot(potsoy)
# potsoy <- ylds$soybean_R  # just taking rainfed commercial potential to start
# potsoy[potsoy > 5] <- round(cellStats(potsoy, mean), 1)
# w <- matrix(nrow = 5, ncol = 5, rep(1, 5 * 5))  # have to filter b/c artifacts
# potsoyf <- focal(potsoy[[1]], w = w, fun = mean, na.rm = TRUE)
# potsoy <- mask(potsoyf, potsoy)  # crop yield
# 
# yield_fix <- function(x = r[[1]], w, na.rm = TRUE) {
#   r <- focal(x, w = w, fun = mean, na.rm = na.rm)
#   rmsk <- (!is.na(r) & is.na(x)) * r
#   x[is.na(x)] <- 0
#   rfix <- x + rmsk  # add back in edge bits
# }
# 
# w <- matrix(3, 3, rep(1, 9))
# potsoyf2 <- yield_fix(potsoy[[1]], w)
# potmaizef2 <- yield_fix(potmaize[[1]], w)

# Disaggregate potential yields, stack, and round
# ylds_1k <- lapply(list(potmaizef2, potsoyf2), function(x) {
#   yld_disagg(x, msk = mgridf)
# })
# ylds_s <- stack(ylds_1k) 
# names(ylds_s) <- c("maize", "soy")
# ylds_s <- round(ylds_s, 1)  
# # plot(ylds_s)
# 
# # data.table them
# cnames <- "ZA"  
# fnm <- full_path(p_production, paste0(cnames, "-potential-yields.csv"))
# DT <- as.data.table(ylds_s, xy = FALSE)[valinds, ]
# # DT[which(is.na(DT))]
# write.table(DT, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
#

### 2016-01-28 Rmd/biodiversity-layer.Rmd
Bunch of stuff, up through creation of figures for presentation, in demo-figures.Rmd

### Weighting in relation to rarity and protectedness
```{r, eval = FALSE}
# Combine protected areas (NPs and GMAs, forest reserves)
nps$code <- ifelse(nps$TYPE == "gma", 2, 1) 
npsr <- rasterize(nps, mgrid, field = "code")  # rasterize NPs/GMAs
npsr[is.na(npsr)] <- 0  
freserves$code <- 3  
freservesr <- rasterize(freserves, mgrid, field = "code")  # forest reserves
freservesr[is.na(freservesr)] <- 0
pas <- npsr + freservesr  # sum them
reclmat <- cbind(0:5, c(0:3, 1, 2))
pas <- reclassify(pas, reclmat)
pas <- mask(pas, mgrid)
# plot(pas)
# plot(freservesr)

@@

writeRaster(bdf, file = full_path(p_bd, "bdprior.tif"), overwrite = TRUE)
vprops <- cbind("var" = c("area", "rarity", "priority"),
                rbind(vtots / nrow(veg_dt), wgti, data.table(t(wgts_prop_np))))
save(vprops, file = full_path(p_bd, "vprops.rda"))

@@

`vprops.rda` - table of veg type area, rarity value, and prioritization value

`bdprior.tif` - biodiversity prioritization grid

>>>>

### Weighting in relation to rarity and protectedness
```{r, eval = FALSE}
# Combine protected areas (NPs and GMAs, forest reserves)
nps$code <- ifelse(nps$TYPE == "gma", 2, 1) 
npsr <- rasterize(nps, mgrid, field = "code")  # rasterize NPs/GMAs
npsr[is.na(npsr)] <- 0  
freserves$code <- 3  
freservesr <- rasterize(freserves, mgrid, field = "code")  # forest reserves
freservesr[is.na(freservesr)] <- 0
pas <- npsr + freservesr  # sum them
reclmat <- cbind(0:5, c(0:3, 1, 2))
pas <- reclassify(pas, reclmat)
pas <- mask(pas, mgrid)
writeRaster(pas, filename = full_path(p_bd, "ZA-pas.tif"))
# plot(pas)
# plot(freservesr)

@@

writeRaster(bdf, file = full_path(p_bd, "cons-priorities.tif"), 
            overwrite = TRUE)
bdprops <- cbind("var" = c("area", "rarity", "priority"),
                 rbind(vtots / nrow(veg_dt), data.table(t(wgtsn)), 
                       data.table(t(wgts_prop_np))))
names(bdprops) <- gsub("veg", "v", names(bdprops))
save(bdprops, file = full_path(p_bd, "ZA-bdprops.rda"))
# load(full_path(p_bd, "ZA-bdprops.rda"))

@@

`ZA-bdprops.rda` - table of veg type area, rarity value, & prioritization value

`cons-priorities.tif` - biodiversity prioritization grid

### 2016-01-28 Rmd/cropland.Rmd
Bunch of stuff, up through creation of figures for presentation, in demo-figures.Rmd

# Planted area and production 2010-2014 for Zambia
muprod <- lapply(c("Maize", "Soybeans"), function(x) {
  fao[Crop == x & var == "prod" & Year %in% c(2008:2015), sum(Value), 
      by = Year]
})
names(muprod) <- c("maize", "soy")
mupa <- lapply(c("Maize", "Soybeans"), function(x) {
  fao[Crop == x & var == "pa" & Year %in% c(2008:2015), sum(Value), 
      by = Year]
})
names(mupa) <- c("maize", "soy")
# mean(muprod$soy[, V1] / mupa$soy[, V1])

>>>>

# Planted area and production 2010-2014 for Zambia
muprod <- lapply(c("Maize", "Soybeans"), function(x) {
  fao[Crop == x & var == "prod" & Year %in% c(2008:2015), sum(Value), 
      by = Year]
})
names(muprod) <- c("maize", "soy")
mupa <- lapply(c("Maize", "Soybeans"), function(x) {
  fao[Crop == x & var == "pa" & Year %in% c(2008:2015), sum(Value), 
      by = Year]
})
names(mupa) <- c("maize", "soy")
# mean(muprod$soy[, V1] / mupa$soy[, V1])
# mean(muprod$maize[, V1] / mupa$maize[, V1])

### 2016-01-28 Rmd/maize-soy-grid.Rmd
Bunch of stuff, up through creation of figures for presentation, in demo-figures.Rmd

# We're going to start by mapping the CV of the soil versus the inter-annual
# variability in the data, just to the 0.25 degree resolution
# read in different grids, start with Maize
crops <- list.dirs(p_crop, full.names = FALSE)[-1]
cv_rasts <- lapply(crops, function(X) {
  # X <- crops[1]
  fnms <- dir(full_path(p_crop, X))
  fl <- unlist(lapply(c("S", "R", "I"), function(x) {
    fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
  }))  # reorder to read subsistence first
  idt <- copy(IDr)
  for(y in 1:length(fl)) { # y <- 2
    print(paste("Processing", fl[y]))
    cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
    cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
    # crss <- cr[YR == 1, ]
    setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
    cr[, WTH := substr(WTH, 1, 4)]
    cr[, c("V1", "TRNO") := NULL]
    setkeyv(cr, c("WTH", "SOL", "Zone"))

@@

Creating 10 km grids for 10th, 50th, 90th percentile yields for each cultivar under each management type (subsistence, rainfed commercial, irrigated). 
```{r, eval=FALSE}
nmso <- c("10", "50", "90")  # naming vector for quantile variables
probs <- c(0.1, 0.5, 0.9)  # quantiles to get
crops <- list.dirs(p_crop, full.names = FALSE)[-1]  # dir names
yld_rasts <- lapply(crops, function(X) {
  # X <- crops[1]
  fnms <- dir(full_path(p_crop, X))
  fl <- unlist(lapply(c("S", "R", "I"), function(x) {
    fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
  }))  # reorder to read subsistence first
  idt <- copy(IDr)
  for(y in 1:length(fl)) { # y <- 1
    print(paste("Processing", fl[y]))
    cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
    cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
    setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
    cr[, WTH := substr(WTH, 1, 4)]
    cr[, c("V1", "TRNO") := NULL]
    setkeyv(cr, c("WTH", "SOL", "Zone"))

@@

# Write out 10 km yield grids to csvs, for safe keeping
lapply(crops, function(x) {
  fnm <- full_path(p_root,
                   paste0("external/data/", x, "_yields_10km.csv"))
  write.csv(yld_rasts[[x]], file = fnm, row.names = FALSE)
})

@@

# ra
w <- matrix(ncol = 3, nrow = 3, rep(1, 9))
yld_rl <- lapply(crops, function(x) {  # x <- crops[[1]]
  yr <- lapply(c("S", "R", "I"), function(y) {  # y <- "S"
    ylds <- yld_rasts[[x]]
    nms <- names(ylds)
    cnms <- nms[grep(paste0("y.*._", y), nms)]
    r <- dt_to_raster(ylds[, c("x", "y", cnms), with = FALSE], CRSobj = gcs)
    rfoc <- sapply(1:nlayers(r), function(i) rfoc <- yield_fix(r[[i]], w = w))
    names(rfoc) <- names(r)
    rs <- stack(rfoc)
    rs
  })
  named_out(yr, c("S", "R", "I"))
})
names(yld_rl) <- crops

@@

## Short-term soybean fix

Potential soybean yields are too low for Zambia, so we will adjust them up with FAO-GAEZ yields

```{r, eval = FALSE}
faonms <- dir(p_fao, full.names = TRUE)
file.rename(faonms, gsub("res02crav6190|yld_package", "", faonms))
faonms <- dir(p_fao, full.names = TRUE)
p_faosoy <- full_path(p_fao, c("soybean_I", "soybean_R", "soybean_S"))
file.rename(faonms, full_path(p_fao, c("soybean_I", "soybean_R", "soybean_S")))
faonms <- dir(p_faosoy, full.names = TRUE)
renm <- paste0(p_faosoy, c("soybean_I", "soybean_R", "soybean_S"), ".tif")
file.rename(faonms[grep("res02_crav6190", faonms)], renm)

@@

# Fix all of them now
soyfix <- lapply(1:3, function(x) {
  ri <- yld_rl$soybean[[x]]
  r1 <- (ri + biasu[2, x]) * (soydiffmsk[[x]] == 0)
  r2 <- ri * (soydiffmsk[[x]] == 1)
  r3 <- r1 + r2
  names(r3) <- names(ri)
  r3
})
names(soyfix) <- names(yld_rl$soybean)
```

```{r, eval = FALSE}
yld_rl2 <- list(yld_rl$maize, soyfix)
names(yld_rl2) <- crops

@@

>>>>

# We're going to start by mapping the CV of the soil versus the inter-annual
# variability in the data, just to the 0.25 degree resolution
# read in different grids, start with Maize
crops <- dir(p_crop, pattern = "maize|soybean", full.names = FALSE)
cv_rasts <- lapply(crops, function(X) {
  # X <- crops[2]
  fnms <- dir(full_path(p_crop, X), pattern = X)
  fl <- unlist(lapply(c("S", "R", "I"), function(x) {
    fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
  }))  # reorder to read subsistence first
  idt <- copy(IDr)

  # idt[, zoneid := rep(0, nrow(idt))]
  # for(i in 1:length(unique(idt$Zone))) {
  #  idt[Zone == unique(idt$Zone)[i], zoneid := i]
  # }

  for(y in 1:length(fl)) { # y <- 3
    print(paste("Processing", fl[y]))
    cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
    cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
    # crss <- cr[YR == 1, ]
    setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
    cr[, WTH := substr(WTH, 1, 4)]
    cr[, c("V1", "TRNO") := NULL]
    setkeyv(cr, c("WTH", "SOL", "Zone"))

@@

Creating 10 km grids for 10th, 50th, 90th percentile yields for each cultivar under each management type (subsistence, rainfed commercial, irrigated). 
```{r, eval=FALSE}
nmso <- c("10", "50", "90")  # naming vector for quantile variables
probs <- c(0.1, 0.5, 0.9)  # quantiles to get
crops <- dir(p_crop, pattern = "maize|soybean", full.names = FALSE)
yld_rasts <- lapply(crops, function(X) {
  # X <- crops[2]
  fnms <- dir(full_path(p_crop, X), pattern = X)
  fl <- unlist(lapply(c("S", "R", "I"), function(x) {
    fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
  }))  # reorder to read subsistence first
  idt <- copy(IDr)
  for(y in 1:length(fl)) { # y <- 1
    print(paste("Processing", fl[y]))
    cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
    cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
    setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
    cr[, WTH := substr(WTH, 1, 4)]
    cr[, c("V1", "TRNO") := NULL]
    setkeyv(cr, c("WTH", "SOL", "Zone"))

@@

# Write out 10 km yield grids to csvs, for safe keeping
lapply(crops, function(x) {
  fnm <- full_path(p_root,
                   paste0("external/data/", x, "_yields_10km.csv"))
  write.csv(yld_rasts[[x]], file = fnm, row.names = FALSE)
})
yld_rasts <- lapply(crops, function(x) {
  fnm <- full_path(p_root,
                   paste0("external/data/", x, "_yields_10km.csv"))
  fread(fnm)
})
names(yld_rasts) <- crops

@@

# ra
w <- matrix(ncol = 3, nrow = 3, rep(1, 9))
yld_rl <- lapply(crops, function(x) {  # x <- crops[[1]]
  yr <- lapply(c("S", "R", "I"), function(y) {  # y <- "S"
    ylds <- yld_rasts[[x]]
    nms <- names(ylds)
    cnms <- nms[grep(paste0("y.*._", y), nms)]
    r <- dt_to_raster(ylds[, c("x", "y", cnms), with = FALSE], 
                      CRSobj = CRS(gcs))
    rfoc <- sapply(1:nlayers(r), function(i) rfoc <- yield_fix(r[[i]], w = w))
    names(rfoc) <- names(r)
    rs <- stack(rfoc)
    rs
  })
  named_out(yr, c("S", "R", "I"))
})
names(yld_rl) <- crops

# cellStats(yld_rl$maize$R[["ymu_R0002"]], mean)
# cellStats(yld_rl$soybean$I[["ymu_I990008"]], mean)
plot(yld_rl$soybean$S[["ymu_S990008"]])
plot(yld_rl$soybean$R[["ymu_R990008"]])
plot(yld_rl$maize$R[["y50_R0003"]])
```

## Yield bias correction
### Soybean fix based on GAEZ data

Potential soybean yields are too low for Zambia, so we will adjust them up with FAO-GAEZ yields. This doesn't seem necessary for maize, because our potential yield averages are at or below the GYGA rainfed yield potentials, per the figures and tables below.

```{r, eval = FALSE} 
cellStats(yld_rl$maize$S[["ymu_S0001"]], mean)
plot(yld_rl$maize$R[["ymu_R0002"]])
plot(yld_rl$maize$S[["ymu_S0002"]])

@@

This is one approach for bias-correction.

```{r, eval = FALSE}
# faonms <- dir(p_fao, full.names = TRUE)
# file.rename(faonms, gsub("res02crav6190|yld_package", "", faonms))
# faonms <- dir(p_fao, full.names = TRUE)
p_faosoy <- full_path(p_fao, c("soybean_I", "soybean_R", "soybean_S"))
file.rename(faonms, full_path(p_fao, c("soybean_I", "soybean_R", "soybean_S")))
faonms <- dir(p_faosoy, full.names = TRUE)
renm <- paste0(p_faosoy, c("soybean_I", "soybean_R", "soybean_S"), ".tif")
file.rename(faonms[grep("res02_crav6190", faonms)], renm)

@@

# Fix all of them now
soyfix <- lapply(1:3, function(x) {
  ri <- yld_rl$soybean[[x]]
  r1 <- (ri + biasu[2, x]) * (soydiffmsk[[x]] == 0)
  r2 <- ri * (soydiffmsk[[x]] == 1)
  r3 <- r1 + r2
  names(r3) <- names(ri)
  r3
})
names(soyfix) <- names(yld_rl$soybean)

Write that out. 
```{r, eval = FALSE}
yld_rl2 <- list(yld_rl$maize, soyfix)
names(yld_rl2) <- crops

@@

But now do the second fix.  Let's adjust yield relative to the actual and baseline and projected growth.

FAO projected growth to 2000-2050 is 69.4% for maize and 219% for soybean. Current mean yields are from FAO 2010-2014 data and USDA

```{r, eval = FALSE}
load(full_path(p_cropland, "currprod.rda"))
load(full_path(p_cropland, "usda-yields.rda"))
load(full_path(p_cropland, "fao-yields.rda"))
cols <- names(usda)[4:9]
usda_ylds <- usda[var == "prod", cols, with = FALSE] / 
  usda[var == "harvarea", cols, with = FALSE]
usda_mu <- round(rowMeans(usda_ylds), 1)

muprod <- lapply(c("Maize", "Soybeans"), function(x) {
  fao[Crop == x & var == "prod" & Year %in% c(2008:2015), sum(Value), 
      by = Year]
})
names(muprod) <- c("maize", "soy")
mupa <- lapply(c("Maize", "Soybeans"), function(x) {
  fao[Crop == x & var == "pa" & Year %in% c(2008:2015), sum(Value), 
      by = Year]
})
names(mupa) <- c("maize", "soy")
fao_mu <- c("maize" = mean(muprod$maize[, V1] / mupa$maize[, V1]), 
            "soy" = mean(muprod$soy[, V1] / mupa$soy[, V1]))

# current yields, mean of USDA and FAO
curr_mu <- round(colMeans(rbind(usda_mu, fao_mu)), 1)  # 2.3 maize, 2.0 soy

# Calculate the mean yields for the country in each management scenario (across 
# cultivar types)
ydiffs <- lapply(crops, function(x) {
  yr <- lapply(c("S", "R", "I"), function(y) {
    r <- yld_rl[[x]][[y]]
    nms <- grep("mu", names(r))
    rmu <- cellStats(calc(r[[nms]], mean), mean)
    rvar <- r / rmu
    names(rvar) <- names(r)
    list("ratio" = rvar, "mu" = rmu)
  })
  names(yr) <- c("S", "R", "I")
  yr
})
names(ydiffs) <- crops

# calculate the mean ratio
yrats <- lapply(1:2, function(x) { # x <- 1
  yr <- lapply(c("S", "R", "I"), function(y) {  # y <- 1
     # inner loop if needed
     X <- crops[x]
     nms <- grep("mu", names(ydiffs[[X]][[y]]$ratio))
     r <- ydiffs[[X]][[y]]$ratio[[nms]]
     rmu <- calc(r, mean)
     rmu 
  })
  names(yr) <- c("S", "R", "I")
  yr
})
names(yrats) <- crops

# Take the DSSAT means and figure out how much they differ relative to their 
# overall mean across management types - use as management specific ratio
dssat_mus <- sapply(ydiffs, function(x) sapply(x, function(y) y$mu))
dssat_adj <- sweep(dssat_mus, MARGIN = 2,  colMeans(dssat_mus), FUN = `/`)

# Bruinsma 2007-2050 growth
bruinsma <- c(maize = 3224 / 1674, soy = 2124 / 1257)

# adjusted DSSAT factors
potmus <- curr_mu * c(bruinsma)  # 4.43 maize, 3.4 soy
potmu_ds <- sweep(dssat_adj, MARGIN = 2,  potmus, FUN = `*`)

# w <- matrix(ncol = 5, nrow = 5, rep(1, 25))
# plot(focal(yrats$maize$S, w), fun = mean, na.rm = TRUE)
# plot(yrats$maize$S * potmu_ds["S", "maize"])
# plot(yrats$maize$R * potmu_ds["R", "maize"])
# plot(yrats$soybean$S * potmu_ds["S", "soybean"])

# finally create adjust yield rasters
ynew <- lapply(crops, function(x) { # x <- 1
  yr <- lapply(c("S", "R", "I"), function(y) {  # y <- 1
     r <- yrats[[x]][[y]] * potmu_ds[y, x]
  })
  names(yr) <- c("S", "R", "I")
  yr
})
names(ynew) <- crops

# writout
for(i in 1:length(ynew)) {
  rl <- ynew[[i]]
  for(j in 1:length(rl)) {
    fnm <- full_path(p_yield, 
                     paste0(crops[i], "_", names(rl)[j], "_10K_FAO.tif"))
    writeRaster(rl[[j]], filename = fnm, overwrite = TRUE)
    layernms <- names(rl[[j]])
    save(layernms, file = gsub("\\.tif", "\\.rda", fnm))
  }
}
